Non-Vertical Cultural Transmission, Assortment, \\and the Evolution of Cooperation

### 2020-10-24 ms/main.tex
Proof that if alpha2>alpha1 then invasion is successful
Add results to tex file

We define the frequencies of the pheno-genotypes $AM$, $BM$, $Am$, and $Bm$ as $\vec{\tilde{p}}= (\tilde{p}_1$, $\tilde{p}_2$, $\tilde{p}_3$, $\tilde{p}_4)$. 
The frequencies of the pheno-genotypes in the next generation are defined by the recursion system 
\begin{equation} \label{eq:next_gen_p_1}
  \begin{aligned}
	% p1
  \bar{w}\tilde{p}_1' & = \tilde{p}_1(\tilde{p}_1+\tilde{p}_3)(1+b-c)(1 - (1-\alpha_1)(\tilde{p}_2 + \tilde{p}_4)T_B) \\
  &+ \tilde{p}_1(\tilde{p}_2 + \tilde{p}_4)(1-c)(1-\alpha_1T_B(\tilde{p}_1 + \tilde{p}_3) - T_B(\tilde{p}_2 + \tilde{p}_4)) \\
  &+ \tilde{p}_2(\tilde{p}_1 + \tilde{p}_3)(1+b)T_A(\tilde{p}_1 + \tilde{p}_3 + \alpha_1(\tilde{p}_2 + \tilde{p}_4)) \\
  &+ \tilde{p}_2(\tilde{p}_2 + \tilde{p}_4)(\tilde{p}_1 + \tilde{p}_3)(1-\alpha_1)T_A 
\\
	% p2
  \bar{w}\tilde{p}_2' & = \tilde{p}_2(\tilde{p}_2+\tilde{p}_4)(1 - (1-\alpha_1)(\tilde{p}_1 + \tilde{p}_3)T_A) \\
  &+ \tilde{p}_2(\tilde{p}_1 + \tilde{p}_3)(1+b)(1-\alpha_1 T_A(\tilde{p}_2 + \tilde{p}_4) - T_A(\tilde{p}_1 + \tilde{p}_3))\\
  &+ \tilde{p}_1(\tilde{p}_2 + \tilde{p}_4)(1-c)(\alpha_1 T_B + (1-\alpha_1)(\tilde{p}_2 + \tilde{p}_4)T_B) \\
  &+ \tilde{p}_1(\tilde{p}_1 + \tilde{p}_3)(1-\alpha_1)(\tilde{p}_2 + \tilde{p}_4)T_B 
\\
	% p3
  \bar{w}\tilde{p}_3' & = \tilde{p}_3(\tilde{p}_1+\tilde{p}_3)(1+b-c)(1 - (1-\alpha_2)(\tilde{p}_2 + \tilde{p}_4)T_B) \\
  &+ \tilde{p}_3(\tilde{p}_2 + \tilde{p}_4)(1-c)(1-\alpha_2T_B(\tilde{p}_1 + \tilde{p}_3) - T_B(\tilde{p}_2 + \tilde{p}_4)) \\
  &+ \tilde{p}_4(\tilde{p}_1 + \tilde{p}_3)(1+b)T_A(\tilde{p}_1 + \tilde{p}_3 + \alpha_2(\tilde{p}_2 + \tilde{p}_4)) \\
  &+ \tilde{p}_4(\tilde{p}_2 + \tilde{p}_4)(\tilde{p}_1 + \tilde{p}_3)(1-\alpha_2)T_A 
\\
	% p4
  \bar{w}\tilde{p}_4' & = \tilde{p}_4(\tilde{p}_2 + \tilde{p}_4)(1 - (1-\alpha_2)(\tilde{p}_1 + \tilde{p}_3)T_A) \\
  &+ \tilde{p}_4(\tilde{p}_1 + \tilde{p}_3)(1+b)(1-\alpha_2T_A(\tilde{p}_2 + \tilde{p}_4)-T_A(\tilde{p}_1 + \tilde{p}_3))\\
  &+ \tilde{p}_3(\tilde{p}_2 + \tilde{p}_4)(1-c)(\alpha_2T_B + (1-\alpha_2)(\tilde{p}_2 + \tilde{p}_4)T_B) \\
  &+ \tilde{p}_3(\tilde{p}_1 + \tilde{p}_3)(1-\alpha_2)(\tilde{p}_2 + \tilde{p}_4)T_B \;,
  \end{aligned}
\end{equation}
%which we can write as 
%\begin{equation}
%\bar{w} \vec{\tilde{p}} = F(\vec{\tilde{p}}),
%\end{equation}
where $\bar{w} = 1 + (b-c)\cdot(\tilde{p}_1 + \tilde{p}_3)$ is the population mean fitness.% and $\vec{\tilde{p}} = (\tilde{p}_1,\tilde{p}_2,\tilde{p}_3,\tilde{p}_4)$.

The local stability of $\vec{\tilde{p}^*} = (\tilde{p}^*, 1-\tilde{p}^*, 0, 0)$, the equilibrium when only allele $M$ is present (see \autoref{eq:vert_hori_equilibrium}), to the introduction of allele $m$ is determined by the linear approximation $\vec{L}^*$ of the transformation in \autoref{eq:next_gen_p_1} near $\vec{\tilde{p}^*}$ (i.e., the Jacobian of the transformation at the equilibrium).
$\vec{L}^*$ is known to have a block structure, with the diagonal blocks occupied by the matrices $\vec{L}^*_{in}$ and $\vec{L}^*_{ex}$.
The latter is the external stability matrix: the linear approximation to the transformation near $\vec{\tilde{p}^*}$ involving only the phono-genotypes $Am$ and $Bm$. Because we assume that $\vec{\tilde{p}^*}$ is internally stable (i.e. locally stable to small perturbations in the frequencies of $AM$ and $BM$), the stability of 
$\vec{\tilde{p}^*}$ is determined by the eigenvalues of the external stability matrix $\vec{L}^*_{ex}$, which is derived from \autoref{eq:next_gen_p_1} as
\begin{equation}
\mathbf{L}^*_{ex} = \frac{1}{\bar{w}^*} \begin{bmatrix}
	 \frac{\partial \tilde{p}_3'}{\partial \tilde{p}_3}(\vec{\tilde{p}}^*) &
	 \frac{\partial \tilde{p}_3'}{\partial \tilde{p}_4}(\vec{\tilde{p}}^*) \\
	 \frac{\partial \tilde{p}_4'}{\partial \tilde{p}_3}(\vec{\tilde{p}}^*) &
	 \frac{\partial \tilde{p}_4'}{\partial \tilde{p}_4} (\vec{\tilde{p}}^*)
\end{bmatrix} \;.
\end{equation}
If the eigenvalues of $\mathbf{L}^*_{ex}$ are less than 1, then the equilibrium $\vec{\tilde{p}}$ is externally stable and allele $m$ cannot invade the population of allele $M$.

\begin{equation} \label{eq:partial_derivatives}
  \begin{aligned}
  &\frac{\partial\tilde{p}_3'}{\partial \tilde{p}_3}(\vec{\tilde{p}}^*) = b(\tilde{p}^*-(1-\alpha_2)T_B(\tilde{p}^*-\tilde{p}^{*2}))+(1-c)(1+\tilde{p}^*T_B-T_B)\\
  &\frac{\partial\tilde{p}_3'}{\partial \tilde{p}_4}(\vec{\tilde{p}}^*) = bT_A\tilde{p}^*(\tilde{p}^*+\alpha_2(1-\tilde{p}^*))+\tilde{p}^*T_A\\
  &\frac{\partial\tilde{p}_4'}{\partial \tilde{p}_3}(\vec{\tilde{p}}^*) = T_B(1-\tilde{p}^*)-cT_B(1-\tilde{p}^*)(\alpha_2+(1-\alpha_2)(1-\tilde{p}^*))\\
  &\frac{\partial\tilde{p}_4'}{\partial \tilde{p}_4}(\vec{\tilde{p}}^*) = 1-\tilde{p}^*T_A+ b\tilde{p}^*(1-\alpha_2T_A(1-\tilde{p}^*)-T_A\tilde{p}^*)\\
  \end{aligned}
\end{equation}

>>>>

We define the frequencies of the pheno-genotypes $AM$, $BM$, $Am$, and $Bm$ as $\vec{\tilde{p}}= (\tilde{p}_1$, $\tilde{p}_2$, $\tilde{p}_3$, $\tilde{p}_4)$. 
The frequencies of the pheno-genotypes in the next generation are defined by the recursion system 
\begin{equation} \label{eq:next_gen_p_1}
  \begin{aligned}
	% p1
  \bar{w}\tilde{p}_1' & = \tilde{p}_1(\tilde{p}_1+\tilde{p}_3)(1+b-c)(1 - (1-\alpha_1)(\tilde{p}_2 + \tilde{p}_4)T_B) \\
  &+ \tilde{p}_1(\tilde{p}_2 + \tilde{p}_4)(1-c)(1-\alpha_1T_B(\tilde{p}_1 + \tilde{p}_3) - T_B(\tilde{p}_2 + \tilde{p}_4)) \\
  &+ \tilde{p}_2(\tilde{p}_1 + \tilde{p}_3)(1+b)T_A(\tilde{p}_1 + \tilde{p}_3 + \alpha_1(\tilde{p}_2 + \tilde{p}_4)) \\
  &+ \tilde{p}_2(\tilde{p}_2 + \tilde{p}_4)(\tilde{p}_1 + \tilde{p}_3)(1-\alpha_1)T_A 
\\
	% p2
  \bar{w}\tilde{p}_2' & = \tilde{p}_2(\tilde{p}_2+\tilde{p}_4)(1 - (1-\alpha_1)(\tilde{p}_1 + \tilde{p}_3)T_A) \\
  &+ \tilde{p}_2(\tilde{p}_1 + \tilde{p}_3)(1+b)(1-\alpha_1 T_A(\tilde{p}_2 + \tilde{p}_4) - T_A(\tilde{p}_1 + \tilde{p}_3))\\
  &+ \tilde{p}_1(\tilde{p}_2 + \tilde{p}_4)(1-c)(\alpha_1 T_B + (1-\alpha_1)(\tilde{p}_2 + \tilde{p}_4)T_B) \\
  &+ \tilde{p}_1(\tilde{p}_1 + \tilde{p}_3)(1+b-c)(1-\alpha_1)(\tilde{p}_2 + \tilde{p}_4)T_B 
\\
	% p3
  \bar{w}\tilde{p}_3' & = \tilde{p}_3(\tilde{p}_1+\tilde{p}_3)(1+b-c)(1 - (1-\alpha_2)(\tilde{p}_2 + \tilde{p}_4)T_B) \\
  &+ \tilde{p}_3(\tilde{p}_2 + \tilde{p}_4)(1-c)(1-\alpha_2T_B(\tilde{p}_1 + \tilde{p}_3) - T_B(\tilde{p}_2 + \tilde{p}_4)) \\
  &+ \tilde{p}_4(\tilde{p}_1 + \tilde{p}_3)(1+b)T_A(\tilde{p}_1 + \tilde{p}_3 + \alpha_2(\tilde{p}_2 + \tilde{p}_4)) \\
  &+ \tilde{p}_4(\tilde{p}_2 + \tilde{p}_4)(\tilde{p}_1 + \tilde{p}_3)(1-\alpha_2)T_A 
\\
	% p4
  \bar{w}\tilde{p}_4' & = \tilde{p}_4(\tilde{p}_2 + \tilde{p}_4)(1 - (1-\alpha_2)(\tilde{p}_1 + \tilde{p}_3)T_A) \\
  &+ \tilde{p}_4(\tilde{p}_1 + \tilde{p}_3)(1+b)(1-\alpha_2T_A(\tilde{p}_2 + \tilde{p}_4)-T_A(\tilde{p}_1 + \tilde{p}_3))\\
  &+ \tilde{p}_3(\tilde{p}_2 + \tilde{p}_4)(1-c)(\alpha_2T_B + (1-\alpha_2)(\tilde{p}_2 + \tilde{p}_4)T_B) \\
  &+ \tilde{p}_3(\tilde{p}_1 + \tilde{p}_3)(1+b-c)(1-\alpha_2)(\tilde{p}_2 + \tilde{p}_4)T_B \;,
  \end{aligned}
\end{equation}
%which we can write as 
%\begin{equation}
%\bar{w} \vec{\tilde{p}} = F(\vec{\tilde{p}}),
%\end{equation}
where $\bar{w} = 1 + (b-c)\cdot(\tilde{p}_1 + \tilde{p}_3)$ is the population mean fitness.% and $\vec{\tilde{p}} = (\tilde{p}_1,\tilde{p}_2,\tilde{p}_3,\tilde{p}_4)$.

The local stability of $\vec{\tilde{p}^*} = (\tilde{p}^*, 1-\tilde{p}^*, 0, 0)$, the equilibrium when only allele $M$ is present (see \autoref{eq:vert_hori_equilibrium}), to the introduction of allele $m$ is determined by the linear approximation $\vec{L}^*$ of the transformation in \autoref{eq:next_gen_p_1} near $\vec{\tilde{p}^*}$ (i.e., the Jacobian of the transformation at the equilibrium).
$\vec{L}^*$ is known to have a block structure, with the diagonal blocks occupied by the matrices $\vec{L}^*_{in}$ and $\vec{L}^*_{ex}$.
The latter is the external stability matrix: the linear approximation to the transformation near $\vec{\tilde{p}^*}$ involving only the phono-genotypes $Am$ and $Bm$. Because we assume that $\vec{\tilde{p}^*}$ is internally stable (i.e. locally stable to small perturbations in the frequencies of $AM$ and $BM$), the stability of 
$\vec{\tilde{p}^*}$ is determined by the eigenvalues of the external stability matrix $\vec{L}^*_{ex}$, which is derived from \autoref{eq:next_gen_p_1} as
\begin{equation}
\mathbf{L}^*_{ex} = \frac{1}{\bar{w}^*} \begin{bmatrix}
	 \frac{\partial \bar{w} \tilde{p}_3'}{\partial \tilde{p}_3}(\vec{\tilde{p}}^*) &
	 \frac{\partial \bar{w}  \tilde{p}_3'}{\partial \tilde{p}_4}(\vec{\tilde{p}}^*) \\
	 \frac{\partial \bar{w}  \tilde{p}_4'}{\partial \tilde{p}_3}(\vec{\tilde{p}}^*) &
	 \frac{\partial \bar{w}  \tilde{p}_4'}{\partial \tilde{p}_4} (\vec{\tilde{p}}^*)
\end{bmatrix} \;.
\end{equation}
If the eigenvalues of $\mathbf{L}^*_{ex}$ are less than 1, then the equilibrium $\vec{\tilde{p}}$ is externally stable and allele $m$ cannot invade the population of allele $M$.

\begin{equation} \label{eq:partial_derivatives}
  \begin{aligned}
  &D_{33} = \frac{\partial\bar{w}\tilde{p}_3'}{\partial \tilde{p}_3}(\vec{\tilde{p}}^*) = b(\tilde{p}^*-(1-\alpha_2)T_B(\tilde{p}^*-\tilde{p}^{*2}))+(1-c)(1+\tilde{p}^*T_B-T_B)\\
  &D_{34} = \frac{\partial\bar{w}\tilde{p}_3'}{\partial \tilde{p}_4}(\vec{\tilde{p}}^*) = bT_A\tilde{p}^*(\tilde{p}^*+\alpha_2(1-\tilde{p}^*))+\tilde{p}^*T_A\\
  &D_{43} =\frac{\partial\bar{w}\tilde{p}_4'}{\partial \tilde{p}_3}(\vec{\tilde{p}}^*) = (1-\tilde{p}^*)(1-c)T_B+b\tilde{p}^*(1-\alpha_2)(1-\tilde{p}^*)T_B\\
  &D_{44} =\frac{\partial\bar{w}\tilde{p}_4'}{\partial \tilde{p}_4}(\vec{\tilde{p}}^*) = 1-\tilde{p}^*T_A+ b\tilde{p}^*(1-\alpha_2T_A(1-\tilde{p}^*)-T_A\tilde{p}^*)\\
  \end{aligned}
\end{equation}
Looking on the greater eigenvalue of $\mathbf{L}^*_{ex}$ $\lambda_1$:
\begin{equation} \label{eq:eigenvalues}
  \lambda_1 = \frac{D_{33} + D_{44} + \sqrt{(D_{33}+D_{44})^2-4*(D_{33}+D_{44}-D_{43}D_{34})}}{2} 
\end{equation}
Using SymPy \emph{SymPy} \citep{Meurer2017}, a Python library for symbolic mathematics we get that $\lambda_1 > 1$ when:
\begin{equation} \label{eq:eigenvalues_greater_than_1_condition}
 \alpha_2 > \frac{b\tilde{p}^*(T_B-T_A) + T_B - T_A  + c (1-T_B)}{b*(T_A(1-\tilde{p}^*) + T_B\tilde{p}^*)}
\end{equation}
Using \autoref{eq:vert_hori_equilibrium} we get that $\lambda_1 > 1$ when:
\begin{equation} \label{eq:eigenvalues_greater_than_1_final_condition}
  \alpha_2 > \alpha_1
 \end{equation}
 % TODO: make this result
